Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-11-09 text/ms.tex
Output and include counts of ups vs. downs
Minor edits

We examined all time series with published explanations of why the black-swan events occurred (Table~\ref{tab:causes-supp}) and focus on seven exemplar populations with high probabilities of heavy tails (Table~\ref{tab:sparks}). The majority (TODO QUANTIFY) of documented events were downward black swans and involved a series of unfortunate events. For example, a shortage of nest sites for the shag long-necked bird (\textit{Phalacrocorax aristotelis}) in the United Kingdom reduced population productivity and made the population more vulnerable when a red tide hit in 1968 \citep{potts1980}. The population experienced a black-swan downswing, but the new availability of nest sites resulted in a rapid population upswing \citep{potts1980}. As another example, an interaction between environmental- and predation-mediated population cycles are thought to have caused a downward black-swan event for a water vole (\textit{Arvicola terrestris}) population \citep{saucy1994}. Other black swans were the result of a sequence of extreme climate events on their own. For example, severe winters in 1929, 1940--1942, and 1962--1963 were associated with black-swan downswings in grey heron (\textit{Ardea cinerea}) abundance in the United Kingdom \citep{stafford1971}. Our analysis finds that the last event was a combination of two black-swan events in a row and it took the population three times longer to recover than predicted \citep{stafford1971}.

(TODO: substitute a parasite example.)

@@

We thank Justin D. Yeakel, TODO, and other members of the Earth to Ocean research group for helpful discussions and comments. We are grateful to the contributors and maintainers of the Global Population Dynamics Database and to Compute Canada's WestGrid high-performance computing resources. Silhouette images were obtained from \texttt{phylopic.org} under a Creative Commons Attribution 3.0 Unported license; sources are listed in the Supporting Information. Funding was provided by a Simon Fraser University Graduate Fellowship (SCA), the Natural Sciences and Engineering Research Council of Canada (NKD, ABC), the Canada Research Chairs Program (NKD).

>>>>

We examined all time series with published explanations of why the black-swan events occurred (Table~\ref{tab:causes-supp}) and focus on seven exemplar populations with high probabilities of heavy tails (Table~\ref{tab:sparks}). The majority (\percBSDown \%) of documented events were downward black swans and involved a series of unfortunate events. For example, a shortage of nest sites for the shag long-necked bird (\textit{Phalacrocorax aristotelis}) in the United Kingdom reduced population productivity and made the population more vulnerable when a red tide hit in 1968 \citep{potts1980}. The population experienced a black-swan downswing, but the new availability of nest sites resulted in a rapid population upswing \citep{potts1980}. As another example, an interaction between environmental- and predation-mediated population cycles are thought to have caused a downward black-swan event for a water vole (\textit{Arvicola terrestris}) population \citep{saucy1994}. Other black swans were the result of a sequence of extreme climate events on their own. For example, severe winters in 1929, 1940--1942, and 1962--1963 were associated with black-swan downswings in grey heron (\textit{Ardea cinerea}) abundance in the United Kingdom \citep{stafford1971}. Our analysis finds that the last event was a combination of two black-swan events in a row and it took the population three times longer to recover than predicted \citep{stafford1971}. \textit{(TODO: substitute a parasite example.)}

@@

We thank Justin D. Yeakel, TODO, and other members of the Earth to Ocean research group for helpful discussions and comments. We are grateful to the contributors and maintainers of the Global Population Dynamics Database and to Compute Canada's WestGrid high-performance computing resources. Silhouette images were obtained from \texttt{phylopic.org} under Creative Commons licenses; sources are listed in the Supporting Information. Funding was provided by a Simon Fraser University Graduate Fellowship (SCA), the Natural Sciences and Engineering Research Council of Canada (NKD, ABC), the Canada Research Chairs Program (NKD).

### 2014-11-09 text/supp.tex
Small edits
Add a line break to make things look better
Minor edits

We fit a multilevel beta regression model to the predicted probability of heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed dynamics. The beta distribution is useful when response data range on a continuous scale between zero and one. We used a logit link function as is typically used in logistic regression. The model was as follows:
\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$ represents the predicted value for population $i$, class $j$, order $k$, and species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$ represents a vector of predictors (such as lifespan) for population $i$ with associated $\beta$ coefficients. The intercepts are allowed to vary from the overall intercept $\alpha$ by taxonomic class ($\alpha^\mathrm{class}_j$), taxonomic order ($\alpha^\mathrm{order}_k$), and species ($\alpha^\mathrm{species}_l$) with standard deviations $\sigma_{\alpha \; \mathrm{class}}$, $\sigma_{\alpha \; \mathrm{order}}$, and $\sigma_{\alpha \; \mathrm{species}}$. Where possible, we also allowed for error distributions around the predictors by incorporating the standard deviation of the posterior samples for the Gompertz parameters $\lambda$, $b$, and $\log \sigma$ around the mean point value as normal distributions (not shown in the above equation).

@@

Many of the silhouette images used in Figs~\ref{fig:nu-coefs}, \ref{fig:correlates} and \ref{fig:posteriors} were obtained from \texttt{phylopic.org} under Creative Commons licenses. We vectorized the salmon in Fig.~\ref{fig:nu-coefs} and Fig.~\ref{fig:correlates} ourselves. The bird in these figures was obtained from \texttt{phylopic.org} under a Creative Commons Attribution 3.0 Unported license with credit to Jean-Raphaël Guillaumin {[}photography{]} and T. Michael Keesey {[}vectorization{]}). The silhouettes in Fig.~\ref{fig:posteriors} were obtained from the following sources:

>>>>

We fit a multilevel beta regression model to the predicted probability of heavy tails, Pr($\nu < 10$), to investigate potential covariates of heavy-tailed dynamics. The beta distribution is useful when response data range on a continuous scale between zero and one \citep{ferrari2004}. We used a logit link function as is typically used in logistic regression. The model was as follows:

\begin{align*}
\mathrm{Pr}(\nu_i < 0.5) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,
\end{align*}
where $A$ and $B$ represent the beta distribution shape parameters; $\mu_i$ represents the predicted value for population $i$, class $j$, order $k$, and species $l$; $\phi_\mathrm{disp}$ represents the dispersion parameter; and $X_i$ represents a vector of predictors (such as lifespan) for population $i$ with associated $\beta$ coefficients. The intercepts are allowed to vary from the overall intercept $\alpha$ by taxonomic class ($\alpha^\mathrm{class}_j$), taxonomic order ($\alpha^\mathrm{order}_k$), and species ($\alpha^\mathrm{species}_l$) with standard deviations $\sigma_{\alpha \; \mathrm{class}}$, $\sigma_{\alpha \; \mathrm{order}}$, and $\sigma_{\alpha \; \mathrm{species}}$. Where possible, we also allowed for error distributions around the predictors by incorporating the standard deviation of the posterior samples for the Gompertz parameters $\lambda$, $b$, and $\log \sigma$ around the mean point value as normal distributions (not shown in the above equation).

@@

Many of the silhouette images used in Figs~\ref{fig:nu-coefs}, \ref{fig:correlates} and \ref{fig:posteriors} were obtained from \texttt{phylopic.org} under Creative Commons licenses. We vectorized the salmon in Fig.~\ref{fig:nu-coefs} and Fig.~\ref{fig:correlates} ourselves. The bird in these figures was obtained from \texttt{phylopic.org} under a Creative Commons Attribution 3.0 Unported license with credit to Jean-Raphaël Guillaumin {[}photography{]} and T. Michael Keesey {[}vectorization{]}). The silhouettes in Fig.~\ref{fig:posteriors} were obtained from the following sources (metadata obtained with the help of the rphylopic R package, \url{https://github.com/sckott/rphylopic}):
